---
title: "Transit Routes"
output:
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
      number_sections: false
    toc_depth: 4
    css: "styles.css"
---

The Route System layer contains the transit routes and stop locations of each transit agency within the region. Each horizon year route is coded over the corresponding horizon year roadway network. Bus routes operating in mixed traffic are coded over roadways, while other modes, such as light rail, which do not operate in mixed traffic, are coded over dedicated guideway links included in the road network layer.  Transit routes use characteristics of the underlying roadway layer in determining speed, travel time, and other attributes.  

## Transit Systems
Transit routes are coded as closely as possible to actual service routes using the model roadway layer . Inbound and outbound trips of a route are coded as separate routes. If the route changes between peak and off-peak (midday), both versions are coded as separate routes. Different route patterns, such as the case where every-other route makes a diversion into a neighborhood are coded as separate routes. The current version of the MRM does not model weekend routes. Night routes are not modeled but are used in estimation of vehicle hours and miles. 

Six separate transit systems are currently operating in the Metrolina region. They are distinguished in the transit route system using the “company” code. The Charlotte Area Transit System (CATS) operates the LYNX light rail, CityLynx streetcar, regional express service, Charlotte express buses and local routes in the Charlotte region. Gastonia Transit operates local routes in Gastonia. Concord-Kannapolis Rider operates routes in Cabarrus County and Salisbury Transit operates routes in Rowan County. Iredell County Transportation Services (ICATS) operates routes in Iredell County.  My Ride Rock Hill operates fixed routes in Rock Hill. 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
library(scales)
library(sf)
library(leaflet)
library(dplyr)

invisible(capture.output({
  mrm_boundary <- st_read("C:/MRM/docs/other/shps/Boundaries/MRMBoundary.shp")
  Transit_22 <- st_read("C:/MRM/docs/other/shps/Transit/TransitNet_2022.shp")
  Transit_25 <- st_read("C:/MRM/docs/other/shps/Transit/TransitNet_2025.shp")
  Transit_35 <- st_read("C:/MRM/docs/other/shps/Transit/TransitNet_2035.shp")
  Transit_45 <- st_read("C:/MRM/docs/other/shps/Transit/TransitNet_2045.shp")
  Transit_55 <- st_read("C:/MRM/docs/other/shps/Transit/TransitNet_2055.shp")

# Ensure both shapefiles are in the same CRS (WGS84)
  mrm_boundary <- st_transform( mrm_boundary, 4326)
  Transit_22 <- st_transform(Transit_22, 4326)
  Transit_25 <- st_transform(Transit_25, 4326)
  Transit_35 <- st_transform(Transit_35, 4326)
  Transit_45 <- st_transform(Transit_45, 4326)
  Transit_55 <- st_transform(Transit_55, 4326)
}))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Create the interactive map
map <- leaflet() %>%
  # Add base tiles
  addTiles() %>%
  
  # Add first shapefile
  addPolygons(
    data =  mrm_boundary,
    color = "black",
    weight = 3,
    opacity = 1,
    fillOpacity = 0,
    highlightOptions = highlightOptions(
      color = "white",
      weight = 3,
      bringToFront = TRUE
    ),
    group = "MRM Boundary",
    labelOptions = labelOptions(noHide = FALSE, direction = "auto")
  ) %>%
  
  
  # Add Transit_2022 shapefile
  addPolylines(
    data = Transit_22,
    color = "blue",
    weight = 2,
    opacity = 0.8,
    highlightOptions = highlightOptions(
      color = "yellow",
      weight = 4,
      bringToFront = TRUE
    ),
    label = ~paste(ROUTE_NAME),
    group = "Transit Network 2022",
    labelOptions = labelOptions(noHide = FALSE, direction = "auto")
  ) %>%
  
  # Add Transit_2025 shapefile
  addPolylines(
    data = Transit_25,
    color = "green",
    weight = 2,
    opacity = 0.8,
    highlightOptions = highlightOptions(
      color = "yellow",
      weight = 4,
      bringToFront = TRUE
    ),
    label = ~paste(ROUTE_NAME),
    group = "Transit Network 2025",
    labelOptions = labelOptions(noHide = FALSE, direction = "auto")
  ) %>%
  
    # Add Transit_2035 shapefile
  addPolylines(
    data = Transit_35,
    color = "orange",
    weight = 2,
    opacity = 0.8,
    highlightOptions = highlightOptions(
      color = "yellow",
      weight = 4,
      bringToFront = TRUE
    ),
    label = ~paste(ROUTE_NAME),
    group = "Transit Network 2035",
    labelOptions = labelOptions(noHide = FALSE, direction = "auto")
  ) %>%
  
    # Add Transit_2045 shapefile
  addPolylines(
    data = Transit_45,
    color = "deeppink",
    weight = 2,
    opacity = 0.8,
    highlightOptions = highlightOptions(
      color = "yellow",
      weight = 4,
      bringToFront = TRUE
    ),
    label = ~paste(ROUTE_NAME),
    group = "Transit Network 2045",
    labelOptions = labelOptions(noHide = FALSE, direction = "auto")
  ) %>%
  
    # Add Transit_2055 shapefile
  addPolylines(
    data = Transit_55,
    color = "purple",
    weight = 2,
    opacity = 0.8,
    highlightOptions = highlightOptions(
      color = "yellow",
      weight = 4,
      bringToFront = TRUE
    ),
    label = ~paste(ROUTE_NAME),
    group = "Transit Network 2055",
    labelOptions = labelOptions(noHide = FALSE, direction = "auto")
  ) %>%
  
  # Layer control
  addLayersControl(
    overlayGroups = c("MRM Boundary","Transit Network 2022","Transit Network 2025","Transit Network 2035","Transit Network 2045","Transit Network 2055"),
    options = layersControlOptions(collapsed = FALSE)
  )

# Display the map
map
```

## Transit Modes 
The MRM has the capability to model 10 modes. Premium modes include light rail (mode 1), streetcar (mode 2), commuter rail (mode 3), and Bus rapid transit (mode 4). The fiscally constrained future year networks contain both light rail and streetcar.   Bus modes include regional express (mode 5), express bus (mode 6), feeder bus (mode 7), local bus (mode 8), and shuttles including the Davidson Shuttle in Davidson (mode 9). The table below provides the fare and type of service modeled for each mode. 

**Mode Fare Table** 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
mode_data <- data.frame(
  "Mode ID" = 1:9,
  "Mode Classification" = c("Premium", "Premium", "Premium", "Premium", "Bus", "Bus", "Bus", "Bus", "Bus"),
  "Mode Description" = c("Light Rail Transit", "Streetcar", "Commuter Rail", "Bus Rapid Transit", 
                         "Regional Express", "Express Bus", "Feeder Bus", "Local Bus", "Downtown Circulator"),
  "Base Year Fare" = c("$1.75", "$1.75", "N/A", "N/A", "$3.50", "$2.40", "N/A", "$1.75", "$0.00"),
  "Future Year Fare" = c("$1.75", "$0.00", "N/A", "N/A", "$3.50", "$2.40", "$0.50", "$1.75", "$0.00")
)

# Create the table with kable
table_modes <- kable(mode_data, col.names = c("Mode ID", "Mode Classification", "Mode Description", 
                                   "Base Year Fare", "Future Year Fare")) %>%
  kable_styling(bootstrap_options = c("bordered", "striped"),
                full_width = TRUE,
                font_size = 14)

# Print the table
table_modes
```


