---
title: "MRM25v2.0"
output:
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
      number_sections: false
    toc_depth: 4
    css: "styles.css"
---

The MRM25v2.0 model was prepared for th2 2026-2035  N.C. Department of Transportation's State Transportation Improvement Program (STIP). The STIP identifies the construction funding and schedule for state transportation projects over a 10-year period. NCDOT updates the STIP approximately every two years.  

As of June 2025, MRM25v2.0 is being reviewed by N.C. Department of Environmental Quality (NC DEQ). The projects represented in the MRM for this model version are described below.

## Model Area

The model area covers a twelve county, bi-state region. In North Carolina, all of Cabarrus, Cleveland, Gaston, Iredell, Lincoln, Mecklenburg, Rowan, Stanly, and Union Counties are included. In South Carolina, all of York County and the northern panhandle of Lancaster County are included. This represents an expansion of the previous model area to include the area of Iredell County north of the Yadkin River. In addition, staff opted to include the southeast corner of Catawba County to capture travel patterns along NC 150 and in the Lake Norman area in general. Figure 1 1 shows the area included in the Metrolina Model.

## Surveys and Studies
MRM22v2.0 builds on the major model update process undertaken with the 2010 Census that included the collection of new travel behavior data as well as building on previously collected data. Tour frequency, destination choice, and time of day models are calibrated based on data collected in the 2012 Household Travel Survey. The mode choice component of the model is calibrated based on data collected in the 2012 Household Travel Survey and the 2013 Transit On-Board Survey.

## Model Years
The official base year validation of MRM22v2.0 is 2018. MTP-based forecast years include 2025, 2035, 2045, and 2050. For conformity, the budget year 2026 was also modeled.

## Highway Networks
The MRM is a regional model. All projects are included where practical and possible. Projects that do not add capacity, are minor intersection improvements, or are small widenings on non-regionally significant roadways may be below the scope of the model and omitted from analysis. All regionally significant projects are included. 

Model projects are referenced by the project index number. A project is selected for a particular model year by index number. Project characteristics (e.g., lanes, facility type, intersection controls) are added to the network before capacities and speeds are estimated. The tables below are grouped by county and sorted by the year the project is to open to traffic. TIP numbers are included where possible.  Some projects cross county and MPO boundaries; these are included at the end of one of the county lists. 


## Project Database

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Load required libraries
library(readxl)
library(DT)
library(dplyr)

# Define a function to render a datatable
render_datatable <- function(data) {
  datatable(
    data,
    filter = "top",
    options = list(
      pageLength = 20,
      autoWidth = TRUE
    ),
    rownames = FALSE
  )
}

data_project <- read_excel("C:/MRM/docs/other/xlsx/MRM25v2_Projects.xlsx") %>%
  mutate(YEAR = as.character(YEAR)) %>%
  select(YEAR, TIP_NO, PROJECT, CATEGORY, TYPE, COUNTY)

# Render the interactive datatable for project list
render_datatable(data_project)

```

## Land Use Statistics

The tables below provide a summary of land use statistics for each of the Metrolina counties.

```{r, echo=FALSE, results='asis'}

# Load required libraries
library(readxl)
library(DT)

# Define a function to render a datatable
render_datatable <- function(data) {
  datatable(
    data,
    filter = "top",
    options = list(
      pageLength = 20,
      autoWidth = TRUE
    ),
    rownames = FALSE
  )
}

# Output plain text for Cabarrus County
cat("**Cabarrus County**  \n")
cat("**Sq. Mi = 364.46**  \n\n")

# Read data from Cabarrus County sheet
data_cabarrus <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Cabarrus3.2")

# Render the interactive datatable for Cabarrus County
render_datatable(data_cabarrus)

# Output plain text for Rowan County
cat("\n**Rowan County**  \n")
cat("**Sq. Mi = 523.68**  \n\n")

# Read data from Rowan County sheet
data_rowan <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Rowan3.3")

# Render the interactive datatable for Rowan County
render_datatable(data_rowan)

# Output plain text for Iredell County
cat("**Iredell County**  \n")
cat("**Sq. Mi = 596.81**  \n\n")

# Read data from Iredell County sheet
data_iredell <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Iredell3.4")

# Render the interactive datatable for Iredell County
render_datatable(data_iredell)

# Output plain text for Mecklenburg County
cat("**Mecklenburg County**  \n")
cat("**Sq. Mi = 545.91**  \n\n")

# Read data from an Excel (.xlsx) file
data_Mecklenburg <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Mecklenburg3.5")

# Render the interactive datatable for Mecklenburg County
render_datatable(data_Mecklenburg)

# Output plain text for Union County
cat("**Union County**  \n")
cat("**Sq. Mi=	639.56**  \n\n")

# Read data from an Excel (.xlsx) file
data_Union <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Union3.6")

# Render the interactive datatable for Union County
render_datatable(data_Union)

# Output plain text for Stanly County
cat("**Stanly County**  \n")
cat("**Sq. Mi=	404.84**  \n\n")

# Read data from an Excel (.xlsx) file
data_Stanly <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Stanly3.7")

# Render the interactive datatable for Stanly County
render_datatable(data_Stanly)

# Output plain text for Cleveland County
cat("**Cleveland County**  \n")
cat("**Sq. Mi=	468.19**  \n\n")

# Read data from an Excel (.xlsx) file
data_Cleveland <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Cleveland3.8")

# Render the interactive datatable for Cleveland County
render_datatable(data_Cleveland)

# Output plain text for Gaston County
cat("**Gaston County**  \n")
cat("**Sq. Mi=	364.10**  \n\n")

# Read data from an Excel (.xlsx) file
data_Gaston <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Gaston3.9")

# Render the interactive datatable for Gaston County
render_datatable(data_Gaston)

# Output plain text for Lincoln County
cat("**Lincoln County**  \n")
cat("**Sq. Mi=	307.03**  \n\n")

# Read data from an Excel (.xlsx) file
data_Lincoln <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Lincoln3.10")

# Render the interactive datatable for Lincoln County
render_datatable(data_Lincoln)

# Output plain text for Lancaster County
cat("**Lancaster County**  \n")
cat("**Sq. Mi=	67.16**  \n\n")

# Read data from an Excel (.xlsx) file
data_Lancaster <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Lancaster3.11")

# Render the interactive datatable for Lancaster County
render_datatable(data_Lancaster)

# Output plain text for York County
cat("**York County**  \n")
cat("**Sq. Mi=	695.84**  \n\n")

# Read data from an Excel (.xlsx) file
data_York <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "York3.12")

# Render the interactive datatable for York County
render_datatable(data_York)

# Output plain text for Catawba County
cat("**Catawba County (partial)**  \n")
cat("**Sq. Mi=	49.25**  \n\n")

# Read data from an Excel (.xlsx) file
data_Catawba <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Catawba3.13")

# Render the interactive datatable for Catawba County
render_datatable(data_Catawba)

```


