---
title: "MRM25v2.0"
output:
  html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
      number_sections: false
    toc_depth: 4
    css: "styles.css"
---

The MRM25v2.0 model was prepared for th2 2026-2035  N.C. Department of Transportation's State Transportation Improvement Program (STIP). The STIP identifies the construction funding and schedule for state transportation projects over a 10-year period. NCDOT updates the STIP approximately every two years.  

As of June 2025, MRM25v2.0 is being reviewed by N.C. Department of Environmental Quality (NC DEQ). The projects represented in the MRM for this model version are described below.

## Model Area

The model area covers a twelve county, bi-state region. In North Carolina, all of Cabarrus, Cleveland, Gaston, Iredell, Lincoln, Mecklenburg, Rowan, Stanly, and Union Counties are included. In South Carolina, all of York County and the northern panhandle of Lancaster County are included. This represents an expansion of the previous model area to include the area of Iredell County north of the Yadkin River. In addition, staff opted to include the southeast corner of Catawba County to capture travel patterns along NC 150 and in the Lake Norman area in general. Figure 1 1 shows the area included in the Metrolina Model.

## Surveys and Studies
MRM22v2.0 builds on the major model update process undertaken with the 2010 Census that included the collection of new travel behavior data as well as building on previously collected data. Tour frequency, destination choice, and time of day models are calibrated based on data collected in the 2012 Household Travel Survey. The mode choice component of the model is calibrated based on data collected in the 2012 Household Travel Survey and the 2013 Transit On-Board Survey.

## Model Years
The official base year validation of MRM22v2.0 is 2018. MTP-based forecast years include 2025, 2035, 2045, and 2050. For conformity, the budget year 2026 was also modeled.

## Highway Networks
The MRM is a regional model. All projects are included where practical and possible. Projects that do not add capacity, are minor intersection improvements, or are small widenings on non-regionally significant roadways may be below the scope of the model and omitted from analysis. All regionally significant projects are included. 

Model projects are referenced by the project index number. A project is selected for a particular model year by index number. Project characteristics (e.g., lanes, facility type, intersection controls) are added to the network before capacities and speeds are estimated. The tables below are grouped by county and sorted by the year the project is to open to traffic. TIP numbers are included where possible.  Some projects cross county and MPO boundaries; these are included at the end of one of the county lists. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# this stuff should be on github not calling local environment
#[Click here to download spreadsheet](https://raw.githubusercontent.com/yourusername/repository/branch/path/to/MRM25v1_Projects.xlsx)

library(DT)
library(dplyr)
library(reactable)
library(htmltools)

# Function to read a fixed-width character field
read_fixed_char <- function(con, length) {
  raw_data <- readBin(con, "raw", n = length)  # Read exact length
  clean_data <- raw_data[raw_data != as.raw(0)]  # Remove null bytes
  rawToChar(clean_data)
}

# Open binary file
con <- file("C:/MRM/MasterNet/Projects.bin", "rb")

# Initialize vectors
indices <- integer()
years <- integer()
tip_nos <- character()
projects <- character()
limits <- character()
categories <- character()
types <- character()
descriptions <- character()
counties <- character()
ec_50mtps <- character()
mtp50_nos <- integer()
projset50s <- character()

# Read records
while (TRUE) {
  # Read integers (4 bytes each)
  index <- readBin(con, "integer", n = 1, size = 4, endian = "little")
  if (length(index) == 0) break  # Stop at EOF
  
  year <- readBin(con, "integer", n = 1, size = 4, endian = "little")
  
  # Read fixed-length character fields
  tip_no <- read_fixed_char(con, 20)
  project <- read_fixed_char(con, 75)
  limit <- read_fixed_char(con, 70)
  category <- read_fixed_char(con, 50)
  type <- read_fixed_char(con, 50)
  description <- read_fixed_char(con, 150)
  county <- read_fixed_char(con, 20)
  ec_50mtp <- read_fixed_char(con, 4)
  
  # Read another integer (4 bytes)
  mtp50_no <- readBin(con, "integer", n = 1, size = 4, endian = "little")
  
  # Read last fixed-length character field
  projset50 <- read_fixed_char(con, 10)
  
  # Append to vectors
  indices <- c(indices, index)
  years <- c(years, year)
  tip_nos <- c(tip_nos, tip_no)
  projects <- c(projects, project)
  limits <- c(limits, limit)
  categories <- c(categories, category)
  types <- c(types, type)
  descriptions <- c(descriptions, description)
  counties <- c(counties, county)
  ec_50mtps <- c(ec_50mtps, ec_50mtp)
  mtp50_nos <- c(mtp50_nos, mtp50_no)
  projset50s <- c(projset50s, projset50)
}

close(con)

# Create data frame with the correct structure
df <- data.frame(
  Index = indices,
  Year = years,
  TIP_NO = tip_nos,
  Project = projects,
  Limits = limits,
  Category = categories,
  Type = types,
  Description = descriptions,
  County = counties,
  EC_50MTP = ec_50mtps,
  MTP50_NO = mtp50_nos,
  PROJSET50 = projset50s,
  stringsAsFactors = FALSE
) %>% 
  select(-EC_50MTP,-MTP50_NO,-PROJSET50) %>% 
  subset(Year <= 2055) %>% 
  arrange(Year, County)

# Convert all character columns to UTF-8 to handle any encoding issues
df[] <- lapply(df, function(x) {
  if (is.character(x)) {
    iconv(x, from = "latin1", to = "UTF-8", sub = "byte")
  } else {
    x
  }
})


```

## Projects Database
```{r, echo=FALSE, warning=FALSE, message=FALSE} 

# Select only the relevant columns
df_filtered <- df %>% 
  select(Year, TIP_NO, Project, Category, Type, County)

# Create a filterable table
reactable(df_filtered, 
          filterable = TRUE,
          searchable = TRUE,
          defaultPageSize = 15,
          striped = TRUE,
          highlight = TRUE,
          bordered = TRUE,
          resizable = TRUE,
          wrap = FALSE,
          defaultColDef = colDef(align = "left"),
         columns = list(
            Year = colDef(filterable = TRUE, width = 100),
            TIP_NO = colDef(filterable = TRUE, width = 150),
            Project = colDef(filterable = TRUE, width = 300, cell = function(value) {
              div(title = value, value)  # Tooltip on hover
            }),
            Category = colDef(filterable = TRUE, width = 200, cell = function(value) {
              div(title = value, value)
            }),
            Type = colDef(filterable = TRUE, width = 200, cell = function(value) {
              div(title = value, value)
            }),
            County = colDef(filterable = TRUE, width = 200, cell = function(value) {
              div(title = value, value)
            })
          ),
          theme = reactableTheme(
            borderColor = "#dee2e6",        
            stripedColor = "#f9f9f9",        
            highlightColor = "lightblue",     
            backgroundColor = "#ffffff",
            headerStyle = list(
              backgroundColor = "#e9ecef",    
              color = "black",            
              fontWeight = 500,              
              textAlign = "left",
              fontSize = "14px",             
              fontFamily = "Segoe UI, -apple-system, system-ui, BlinkMacSystemFont, Roboto, sans-serif"
            ),
            cellStyle = list(
              padding = "12px 8px",          
              fontSize = "14px",             
              fontFamily = "Segoe UI, -apple-system, system-ui, BlinkMacSystemFont, Roboto, sans-serif",
              borderBottom = "1px solid #dee2e6"
            ),
            inputStyle = list(
              backgroundColor = "#ffffff",
              borderColor = "#ced4da",      
              color = "#495057",            
              fontSize = "14px"
            )
          ),
          width = "100%" 
)


```

## Land Use Statistics

The tables below provide a summary of land use statistics for each of the Metrolina counties.

```{r, echo=FALSE, results='asis'}

# Load required libraries
library(readxl)

# Output plain text above the table (NOT markdown headers)
cat("**Cabarrus County**  \n")
cat("**Sq. Mi = 364.46**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Cabarrus3.2")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Rowan County**  \n")
cat("**Sq. Mi = 523.68**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Rowan3.3")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```


```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Iredell County**  \n")
cat("**Sq. Mi = 596.81**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Iredell3.4")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Mecklenburg County**  \n")
cat("**Sq. Mi = 545.91**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Mecklenburg3.5")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Union County**  \n")
cat("**Sq. Mi=	639.56**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Union3.6")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Stanly County**  \n")
cat("**Sq. Mi=	404.84**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Stanly3.7")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Cleveland County**  \n")
cat("**Sq. Mi=	468.19**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Cleveland3.8")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Gaston County**  \n")
cat("**Sq. Mi=	364.10**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Gaston3.9")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Lincoln County**  \n")
cat("**Sq. Mi=	307.03**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Lincoln3.10")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Lancaster County**  \n")
cat("**Sq. Mi=	67.16**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Lancaster3.11")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**York County**  \n")
cat("**Sq. Mi=	695.84**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "York3.12")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

```{r, echo=FALSE, results='asis'}

# Output plain text above the table (NOT markdown headers)
cat("**Catawba County (partial)**  \n")
cat("**Sq. Mi=	695.84**  \n\n")

# Read data from an Excel (.xlsx) file
data <- read_excel("C:/MRM/docs/other/xlsx/LandUseStat.xlsx", sheet = "Catawba3.13")

# Render the interactive datatable
datatable(
  data,
  filter = "top",
  options = list(
    pageLength = 20,
    autoWidth = TRUE
  ),
  rownames = FALSE
)
```

